---
title: "Advanced: Simulating batch effects"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteEncoding{UTF-8}
  %\VignetteIndexEntry{Advanced: Simulating batch effects}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

<!-- github markdown built using 
rmarkdown::render("vignettes/advanced_simulating_batch_effects.Rmd", output_format = rmarkdown::github_document())
-->

An easy way of simulating batch effects is by performing multiple dyngen runs, but with different kinetics.
dyngen samples the GRN kinetics randomly from a set of predefined distributions (See `?generate_kinetics`).


## Create underlying gene regulatory network
First, create the 'common' part of the dyngen simulation as follows. This will 
ensure that the gene regulatory network-part of the simulation is exactly the same.
```{r init, message=FALSE, fig.width = 10, fig.height = 6}
library(tidyverse)
library(dyngen)

set.seed(1)

model_common <-
  initialise_model(
  backbone = backbone_bifurcating(),
  num_cells = 1000,
  num_tfs = 50,
  num_targets = 250,
  num_hks = 250,
  simulation_params = simulation_default(
    census_interval = 10, 
    ssa_algorithm = ssa_etl(tau = 300 / 3600),
    experiment_params = simulation_type_wild_type(num_simulations = 100)
  )
) %>%
  generate_tf_network() %>%
  generate_feature_network()
```

## Simulate multiple times with different kinetics

Now you can run the simulation multiple times.
Note that for each separate run, the `generate_kinetics()` step ensures that the 
thermodynamic parameters will be different (yet drawn from the same distribution).
Some values for the kinetics of the main transcription factors are not changed, 
to ensure that the desired dynamic process is obtained.
```{r simulation} 
model_a <- model_common %>%
  generate_kinetics() %>%
  generate_gold_standard() %>%
  generate_cells()

model_b <- model_common %>%
  generate_kinetics() %>%
  generate_gold_standard() %>%
  generate_cells()
```

The differences if kinetics parameters can be visualised as follows.
```{r compare_kinetics}
params_a <- 
  dyngen:::.kinetics_extract_parameters_as_df(model_a$feature_info, model_a$feature_network) %>% 
  select(id, group = param, model_a = value)

params_b <- 
  dyngen:::.kinetics_extract_parameters_as_df(model_b$feature_info, model_b$feature_network) %>% 
  select(id, group = param, model_b = value)

params_ab <-
  inner_join(params_a, params_b, by = c("id", "group"))

ggplot(params_ab) + 
  geom_point(aes(model_a, model_b, colour = group)) + 
  facet_wrap(~group, scales = "free") +
  theme_bw()
```

## Combine outputs and visualise
Finally, combine the simulations as follows.
```{r combine}
model_ab <-
  combine_models(list(left = model_a, right = model_b)) %>% 
  generate_experiment()
```

Show a dimensionality reduction.
```{r plot, fig.width = 10, fig.height = 6}
plot_simulations(model_ab)
plot_gold_mappings(model_ab, do_facet = FALSE)
```

Visualise the dataset using dyno.
```{r dyno, fig.width = 10, fig.height = 6}
library(dyno)
dataset <- as_dyno(model_ab)
plot_dimred(dataset)
plot_heatmap(dataset, features_oi = 50)
```

## Additional thoughts
It's definitely possible to think of different ways of simulating batch effects. 
For instance, you could implement a different strategy for changing the kinetics parameters
between runs. You could also change the gene regulatory network itself, e.g. by changing the
effect of some of the edges, or changing the targets of certain regulators. 

If you come up with something, feel free to let us know on GitHub!