---
title: "Advanced: Simulating a knockout experiment"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteEncoding{UTF-8}
  %\VignetteIndexEntry{Advanced: Simulating a knockout experiment}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

<!-- github markdown built using 
rmarkdown::render("vignettes/advanced_simulating_knockouts.Rmd", output_format = rmarkdown::github_document())
-->

dyngen supports simulating knockout (or knockdown) experiments. In this experiment, dyngen is run twice (similar to the vignette on simulating batch effects), 
once for the wild type (WT), and once for the knockout (KO).

## Initialising the experiment
First, create the 'common' part of the dyngen simulation as follows. This will 
ensure that the gene regulatory network-part of the simulation is exactly the same.
```{r init, message=FALSE, fig.width = 8, fig.height = 6}
library(tidyverse)
library(dyngen)

set.seed(1)

backbone <- backbone_bifurcating()

model_common <-
  initialise_model(
  backbone = backbone,
  num_cells = 1000,
  num_tfs = nrow(backbone$module_info),
  num_targets = 250,
  num_hks = 250,
  simulation_params = simulation_default(
    census_interval = 10, 
    ssa_algorithm = ssa_etl(tau = 300 / 3600),
    experiment_params = simulation_type_wild_type(num_simulations = 100)
  )
) %>%
  generate_tf_network() %>%
  generate_feature_network() %>% 
  generate_kinetics() %>%
  generate_gold_standard()
```

## Simulate wild type

Simulating the wild type is easy, as it is the standard dyngen way of simulating cells.

```{r wildtype, fig.width = 8, fig.height = 6} 
model_wt <- model_common %>%
  generate_cells()

plot_gold_mappings(model_wt, do_facet = FALSE)
```

## Simulate knockout

To simulate a gene knockout, the experiment type of the simulation is changed from wild type (see first code block) to knockdown. 
The `simulation_type_knockdown()` function is actually designed to perform single-cell knockdowns of 1 to 5 random genes, but 
this behaviour can be overridden by setting the following parameters:

 * `genes`: list of gene ids to be knocked out,
 * `num_genes`: number of genes to be knocked out per cell (sampled randomly),
 * `multiplier`: the fraction of transcription rate after KO (0 → no transcription, 1 → no KO).
 
In this case, looking at the module network of the bifurcation model it is possible
to see that the B3 gene module is important for diversification of cells towards one particular
cell lineage (to verify, one can draw out `model_common$module_network` on paper).

```{r visnet}
plot_backbone_modulenet(model_common)
b3_genes <- model_common$feature_info %>% filter(module_id == "B3") %>% pull(feature_id)
```

By knocking out the gene "`r cat(b3_genes)`", one of the two branches will be inhibited.
```{r knockout}
model_ko <- model_common
model_ko$simulation_params$experiment_params <- simulation_type_knockdown(
  num_simulations = 100L,
  timepoint = 0, 
  genes = b3_genes,
  num_genes = length(b3_genes),
  multiplier = 0
)
```

Running the simulation will generate a warning stating that some of the gold standard edges 
were not covered by any of the simulations, but that's to be expected since we knocked it out. 
We can see that indeed no simulation managed to enter the second branch (with end state D).
```{r kosim, fig.width = 8, fig.height = 6}
model_ko <- model_ko %>%
  generate_cells()

plot_gold_mappings(model_ko, do_facet = FALSE)
```

## Combine outputs and visualise
We can combine both simulations into one simulation and visualise the output.
```{r combine, fig.width = 8, fig.height = 6}
model_comb <-
  combine_models(list(WT = model_wt, KO = model_ko)) %>% 
  generate_experiment()

plot_simulations(model_comb)
plot_gold_mappings(model_comb, do_facet = FALSE)
```

Visualise the dataset using dyno.
```{r dyno, fig.width = 10, fig.height = 7}
library(dyno)
dataset <- as_dyno(model_comb)
plot_dimred(dataset)
plot_heatmap(dataset, features_oi = 50)
```
