---
title: "Getting started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

<!-- github markdown built using 
rmarkdown::render("vignettes/getting_started.Rmd", output_format = rmarkdown::github_document())
-->

This vignette demonstrates the basics of running a dyngen simulation. 
If you haven't done so already, first check out the installation instructions 
in the README.

## Step 1: Define backbone and other parameters
A dyngen simulation can be started by providing a backbone to the `initialise_model()` function.
The backbone of a `dyngen` model is what determines the overall dynamic process 
that a cell will undergo during a simulation. It consists of a set of gene modules, which regulate
eachother in such a way that expression of certain genes change over time in a specific manner.

```{r init, message=FALSE, fig.width = 10, fig.height = 6}
library(tidyverse)
library(dyngen)

set.seed(10)
model <- 
  initialise_model(
    num_tfs = 12,
    num_targets = 30,
    num_hks = 15,
    backbone = backbone_bifurcating(),
    verbose = TRUE
  )

plot_backbone_statenet(model)
plot_backbone_modulenet(model)
```

For backbones with all different sorts of topologies, check `list_backbones()`:
```{r list_backbones}
names(list_backbones())
```

## Step 2: Generate transcription factors (TFs)
Each gene module consists of a set of transcription factors.
These can be generated and visualised as follows.
```{r tf_network}
model <- generate_tf_network(model)
plot_feature_network(model, show_targets = FALSE)
```

## Step 3: Sample target genes and housekeeping genes (HKs) 
Next, target genes and housekeeping genes are added to the network by
sampling a gold standard gene regulatory network using the Page Rank algorithm.
Target genes are regulated by TFs or other target genes, while HKs are only regulated
by themselves.
```{r target_network}
model <- generate_feature_network(model)
plot_feature_network(model)
plot_feature_network(model, show_hks = TRUE)
```

## Step 4: Generate kinetics
Note that the target network does not show the effect of some interactions, 
because these are generated along with other kinetics parameters of the 
SSA simulation.

```{r ssa}
model <- generate_kinetics(model)
plot_feature_network(model)
plot_feature_network(model, show_hks = TRUE)
```

## Step 5: Simulate gold standard
The gold standard is simulated by enabling certain parts of 
the module network and performing ODE simulations. The gold standard
are visualised by performing a dimensionality reduction on the 
mRNA expression values.
```{r gold_standard}
model <- generate_gold_standard(model)
plot_gold_simulations(model) + scale_colour_brewer(palette = "Dark2")
```

The expression of the modules (average of TFs) can be visualised as follows.
```{r gold_pt, fig.width=10, fig.height=10}
plot_gold_expression(model, what = "mol_mrna") # mrna
plot_gold_expression(model, label_changing = FALSE) # premrna, mrna, and protein
```

## Step 6: Simulate cells.
Cells are simulated by running SSA simulations. The simulations are again
using dimensionality reduction.
```{r simulations}
model <- generate_cells(model)
plot_simulations(model)
```

The gold standard can be overlayed on top of the simulations.
```{r overlay}
plot_gold_simulations(model) + scale_colour_brewer(palette = "Dark2")
```

We can check how each segment of a simulation is mapped to the gold standard.
```{r compare}
plot_gold_mappings(model, do_facet = FALSE) + scale_colour_brewer(palette = "Dark2")
```

The expression of the modules (average of TFs) of a single simulation can be visualised as follows.
```{r expression_sim}
plot_simulation_expression(model, 1:4, what = "mol_mrna")
```

## Step 7: Experiment emulation
Effects from performing a single-cell RNA-seq experiment can be emulated as follows.
```{r experiment}
model <- generate_experiment(model)
```

## Step 8: Convert to a dyno object
Converting the dyngen to a dyno object allows you to visualise the dataset
using the `dynplot` functions, or infer trajectories using `dynmethods`.
```{r wrap}
dataset <- as_dyno(model)
```

### Visualise with `dynplot`
```{r dynplot}
library(dynplot)
plot_dimred(dataset)
plot_graph(dataset)
```

### Infer trajectory on expression data
```{r dyno}
library(dyno)
pred <- infer_trajectory(dataset, ti_slingshot())
plot_dimred(pred)
```

### Save output to file
You can save the output of the model and/or dataset to a file as follows.
```r
write_rds(model, "model.rds", compress = "gz")
write_rds(dataset, "dataset.rds", compress = "gz")
```

## Step 8 alternative: Convert to an anndata object
dyngen (>= 0.4.1) allows converting the output to an `anndata` object as well.
Check out the [anndata documentation](https://cran.r-project.org/package=anndata) 
on how to install anndata for R.

```r
library(anndata)
ad <- as_anndata(model)
ad$write_h5ad("dataset.h5ad")
```

# One-shot function

`dyngen` also provides a one-shot function for running 
all of the steps all at once and producing plots.
```{r oneshot_run, message=FALSE, fig.width=20, fig.height=20}
set.seed(1)
config <- 
  initialise_model(
    num_tfs = 12,
    num_targets = 30,
    num_hks = 15,
    backbone = backbone_bifurcating_converging(),
    verbose = FALSE
  )

out <- generate_dataset(
  config,
  make_plots = TRUE
)

dataset <- out$dataset
model <- out$model
print(out$plot)
```

`dataset` and `model` can be used in much the same way as before.

```{r oneshot_plot}
plot_dimred(dataset)
plot_graph(dataset)
pred <- infer_trajectory(dataset, ti_slingshot(), verbose = FALSE)
plot_dimred(pred)
```
