---
title: "Showcase different backbones"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{showcase_backbones}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

<!-- github markdown built using 
rmarkdown::render("vignettes/showcase_backbones.Rmd", output_format = "github_document")
-->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 20,
  fig.height = 15
)
```

```{r setup}
library(dyngen)
```

dyngen supports different trajectory topologies, such as bifurcating and cyclic. You can find a full list of backbones using `?list_backbones`. This vignette will showcase each of them individually.

# Linear

```{r linear}
backbone <- backbone_linear()

init <- initialise_model(
  backbone = backbone,
  num_cells = 500,
  num_tfs = 100,
  num_targets = 50,
  num_hks = 25,
  simulation_params = simulation_default(census_interval = 10, ssa_algorithm = ssa_etl(tau = 300 / 3600)),
  download_cache_dir = "~/.cache/dyngen/",
  num_cores = 7
)
out <- generate_dataset(init, make_plots = TRUE)
out$plot
```


# Bifurcating
```{r bifurcating}
backbone <- backbone_bifurcating()

init <- initialise_model(
  backbone = backbone,
  num_cells = 500,
  num_tfs = 100,
  num_targets = 50,
  num_hks = 25,
  simulation_params = simulation_default(census_interval = 10, ssa_algorithm = ssa_etl(tau = 300 / 3600)),
  download_cache_dir = "~/.cache/dyngen/",
  num_cores = 7
)
out <- generate_dataset(init, make_plots = TRUE)
out$plot
```

# Bifurcating converging
```{r bifurcating_converging}
backbone <- backbone_bifurcating_converging()

init <- initialise_model(
  backbone = backbone,
  num_cells = 500,
  num_tfs = 100,
  num_targets = 50,
  num_hks = 25,
  simulation_params = simulation_default(census_interval = 10, ssa_algorithm = ssa_etl(tau = 300 / 3600)),
  download_cache_dir = "~/.cache/dyngen/",
  num_cores = 7
)
out <- generate_dataset(init, make_plots = TRUE)
out$plot
```

# Bifurcating cycle
```{r bifurcating_cycle}
backbone <- backbone_bifurcating_cycle()

init <- initialise_model(
  backbone = backbone,
  num_cells = 500,
  num_tfs = 100,
  num_targets = 50,
  num_hks = 25,
  simulation_params = simulation_default(census_interval = 10, ssa_algorithm = ssa_etl(tau = 300 / 3600)),
  download_cache_dir = "~/.cache/dyngen/",
  num_cores = 7
)
out <- generate_dataset(init, make_plots = TRUE)
out$plot
```

# Bifurcating loop
```{r bifurcating_loop}
backbone <- backbone_bifurcating_loop()

init <- initialise_model(
  backbone = backbone,
  num_cells = 500,
  num_tfs = 100,
  num_targets = 50,
  num_hks = 25,
  simulation_params = simulation_default(census_interval = 10, ssa_algorithm = ssa_etl(tau = 300 / 3600)),
  download_cache_dir = "~/.cache/dyngen/",
  num_cores = 7
)
out <- generate_dataset(init, make_plots = TRUE)
out$plot
```

# Binary tree
```{r binary_tree}
backbone <- backbone_binary_tree(
  num_modifications = 2
)

init <- initialise_model(
  backbone = backbone,
  num_cells = 500,
  num_tfs = 100,
  num_targets = 50,
  num_hks = 25,
  simulation_params = simulation_default(census_interval = 10, ssa_algorithm = ssa_etl(tau = 300 / 3600)),
  download_cache_dir = "~/.cache/dyngen/",
  num_cores = 7
)
out <- generate_dataset(init, make_plots = TRUE)
out$plot
```

# Branching
```{r branching}
backbone <- backbone_branching(
  num_modifications = 2,
  min_degree = 3,
  max_degree = 3
)

init <- initialise_model(
  backbone = backbone,
  num_cells = 500,
  num_tfs = 100,
  num_targets = 50,
  num_hks = 25,
  simulation_params = simulation_default(census_interval = 10, ssa_algorithm = ssa_etl(tau = 300 / 3600)),
  download_cache_dir = "~/.cache/dyngen/",
  num_cores = 7
)
out <- generate_dataset(init, make_plots = TRUE)
out$plot
```

# Consecutive bifurcating
```{r consecutive_bifurcating}
backbone <- backbone_consecutive_bifurcating()

init <- initialise_model(
  backbone = backbone,
  num_cells = 500,
  num_tfs = 100,
  num_targets = 50,
  num_hks = 25,
  simulation_params = simulation_default(census_interval = 10, ssa_algorithm = ssa_etl(tau = 300 / 3600)),
  download_cache_dir = "~/.cache/dyngen/",
  num_cores = 7
)
out <- generate_dataset(init, make_plots = TRUE)
out$plot
```

# Trifurcating
```{r trifurcating}
backbone <- backbone_trifurcating()

init <- initialise_model(
  backbone = backbone,
  num_cells = 500,
  num_tfs = 100,
  num_targets = 50,
  num_hks = 25,
  download_cache_dir = "~/.cache/dyngen/",
  num_cores = 7
)
out <- generate_dataset(init, make_plots = TRUE)
out$plot
```

# Converging
```{r converging}
backbone <- backbone_converging()

init <- initialise_model(
  backbone = backbone,
  num_cells = 500,
  num_tfs = 100,
  num_targets = 50,
  num_hks = 25,
  simulation_params = simulation_default(census_interval = 10, ssa_algorithm = ssa_etl(tau = 300 / 3600)),
  download_cache_dir = "~/.cache/dyngen/",
  num_cores = 7
)
out <- generate_dataset(init, make_plots = TRUE)
out$plot
```

# Cycle
```{r cycle}
backbone <- backbone_cycle()

init <- initialise_model(
  backbone = backbone,
  num_cells = 500,
  num_tfs = 100,
  num_targets = 50,
  num_hks = 25,
  simulation_params = simulation_default(census_interval = 10, ssa_algorithm = ssa_etl(tau = 300 / 3600)),
  download_cache_dir = "~/.cache/dyngen/",
  num_cores = 7
)
out <- generate_dataset(init, make_plots = TRUE)
out$plot
```

# Disconnected
```{r disconnected}
backbone <- backbone_disconnected()

init <- initialise_model(
  backbone = backbone,
  num_cells = 500,
  num_tfs = 100,
  num_targets = 50,
  num_hks = 25,
  simulation_params = simulation_default(census_interval = 10, ssa_algorithm = ssa_etl(tau = 300 / 3600)),
  download_cache_dir = "~/.cache/dyngen/",
  num_cores = 7
)
out <- generate_dataset(init, make_plots = TRUE)
out$plot
```

# Linear simple
```{r linear_simple}
backbone <- backbone_linear_simple()

init <- initialise_model(
  backbone = backbone,
  num_cells = 500,
  num_tfs = 100,
  num_targets = 50,
  num_hks = 25,
  download_cache_dir = "~/.cache/dyngen/",
  num_cores = 7
)
out <- generate_dataset(init, make_plots = TRUE)
out$plot
```

# Cycle simple
```{r cycle_simple}
backbone <- backbone_cycle_simple()

init <- initialise_model(
  backbone = backbone,
  num_cells = 500,
  num_tfs = 100,
  num_targets = 50,
  num_hks = 25,
  download_cache_dir = "~/.cache/dyngen/",
  num_cores = 7
)
out <- generate_dataset(init, make_plots = TRUE)
out$plot
```